Incentivizing Exploration by Heterogeneous Users

### 2018-06-05 introduction.tex
significant edits throughout

\section{Introduction}

@@

To effect an outcome close to societally optimal,
it is necessary to provide exploration incentives to the individual users.
This was noted in two recent lines of work:
\citet{kremer2014implementing}
and \citet{mansour2015bayesian,mansour2016bayesian}
assume that the site (also called the \emph{principal}) has an
informational advantage in being the only one to observe the results
of past arm pulls.
(Such an assumption applies, for instance, to route recommendations in
driving.)
The principal can exploit her%
\footnote{We use male pronouns to refer to users and female pronouns
  to refer to the principal.}
advantage and make recommendations to the individual agents that 
are in their best interest to follow.
\citet{frazier2014incentivizing} and 
\citet{han2015incentivizing} instead assume that the results of all
past arm pulls are publicly observable
(such as reviews on an online retail site).
They instead suppose the principal can offer payments to users as
a reward for pulling particular arms.
We follow the model of \citet{frazier2014incentivizing} and
\citet{han2015incentivizing} 
and consider a multi-armed bandit model in which the principal can
offer payments to the users for pulling particular arms.

@@

Heterogeneity presents both a challenge and an opportunity.
On the one hand, unobserved heterogeneity hides critical
information about an agent's preferences from the principal.
On the other hand, heterogeneity also presents her an opportunity,
through the possibility of ``free exploration.''
Even when left unincentivized, agents will
play a variety of arms, revealing information about their attributes.
This stands in sharp contrast to the case of homogeneous preferences,
where unincentivized agents will herd onto a single apparently best arm;
thus, effecting essentially any exploration at all requires incentives.

@@

Toward that end, we model our setting as follows.
(We describe our model at a high level here,
with formal definitions given in Section~\ref{sec:prob}.)
The \ARMNUM arms and users (or \emph{agents})
are characterized by payoff-relevant
\emph{attribute} (or \emph{feature}) vectors.
Arms' attributes are a priori unknown,
and agents' attributes are drawn from a known distribution.
An agent's reward from pulling an arm is the inner product of his
vector with the arm's vector (plus noise).
When an arm is pulled, a noisy version of its attribute vector is
observed by everyone.
\pfedit{This models full-text product reviews on websites like Amazon, or ratings of ``service'', ``value'', and other restaurant attributes on websites like Tripadvisor.  }
Agents are myopic and will pull the arm whose expected attribute
vector (based on past noisy observations) maximizes their reward.
\pfedit{Agents' ability to effectively base decisions on all past observations models platforms' abilities to aggregate and effectively display past feedback, through simple averages of ratings and automatic summarization of full-text reviews \citep{wang2010product,liu2012movie,abulaish2009feature}.  While platforms can manipulate the display of user data, most have an incentive to be seen as truthful.}
The principal can incentivize agents to pull particular arms by
offering rewards for the specific arm.
The principal's goal is to keep the cumulative regret across all
agents small, while incurring only small total payments.
Our main theorem can be stated informally as follows:

@@

\begin{theorem} \label{thm:main-intro}
Assume that for each arm, at least a constant fraction \MinProb
of the population likes this arm best,
and let \TieDensity be a measure of the ``density of near-ties''
between agents' arm preferences
(in a sense made precise in Section~\ref{subsec:discrete}).
There is a policy that achieves expected 
cumulative regret $O (\ARMNUM \e^{2/\MinProb} + \ARMNUM \log^3(T))$,
using expected cumulative payments of $O(\ARMNUM^2 \e^{2/\MinProb})$.
In particular, when agents who are close to tied between two arms have measure $0$,
both the expected regret and expected payment are bounded by constants
(with respect to $T$). 
\end{theorem}

>>>>

\section{Introduction} \label{sec:introduction}

@@

To effect an outcome close to societally optimal,
it is necessary to provide exploration incentives to the individual users.
This was noted in two recent lines of work:
\citet{kremer2014implementing}
and
\citet{mansour2015bayesian,mansour2016bayesian,mansour2018competition}
(see \cite{slivkins:asymmetry} for a high-level overview)
assume that the site (also called the \emph{principal}) has an
informational advantage in being the only one to observe the results
of past arm pulls.
(Such an assumption applies, for instance, to route recommendations in
driving.)
The principal can exploit her%
\footnote{We use male pronouns to refer to users and female pronouns
  to refer to the principal.}
advantage and make recommendations to the individual agents that 
are in their best interest to follow.
\citet{frazier2014incentivizing} and 
\citet{han2015incentivizing} instead assume that the results of all
past arm pulls are publicly observable
(such as reviews on an online retail site).
They instead suppose that the principal can offer payments to users as
a reward for pulling particular arms.
We follow the model of \citet{frazier2014incentivizing} and
\citet{han2015incentivizing} 
and consider a multi-armed bandit model in which the principal can
offer payments to the users for pulling particular arms.

@@

Heterogeneity presents both a challenge and an opportunity.
On the one hand, unobserved heterogeneity hides critical
information about an agent's preferences from the principal.
On the other hand, heterogeneity also presents her an opportunity,
through the possibility of ``free exploration.''
Even when left unincentivized, agents will
play a variety of arms, revealing information about the arms' attributes.
This stands in sharp contrast to the case of homogeneous preferences,
where unincentivized agents will herd onto a single apparently best arm;
thus, effecting essentially any exploration at all requires incentives.

@@

Toward that end, we model our setting as follows.
(We describe our model at a high level here,
with formal definitions given in Section~\ref{sec:prob}.)
The \ARMNUM arms and users (or \emph{agents})
are characterized by payoff-relevant
\emph{attribute} (or \emph{feature}) vectors.
Arms' attributes are a priori unknown,
and agents' attributes are drawn from a known distribution.
An agent's reward from pulling an arm is the inner product of his
vector with the arm's vector (plus noise).
When an arm is pulled, a noisy version of its attribute vector is
observed by everyone.
This models full-text product reviews on websites like Amazon,
or ratings of ``service'', ``value'',
and other restaurant attributes on websites like Tripadvisor.  
Agents are myopic and will pull the arm whose expected attribute
vector (based on past noisy observations) maximizes their reward.
\dkedit{This assumption requires agents to have access to aggregate
information about the feature vector estimates of all (relevant) arms. 
Such access is facilitated by the platforms' ability to aggregate and
effectively display past feedback, e.g., by displaying simple averages
of ratings and using automatic summarization of full-text reviews 
\citep{wang2010product,liu2012movie,abulaish2009feature}.
While platforms could in principle manipulate the display of such
data, most have an incentive to be seen as truthful.}
%\pfedit{Agents' ability to effectively base decisions on all past observations models platforms' abilities to aggregate and effectively display past feedback, through simple averages of ratings and automatic summarization of full-text reviews \citep{wang2010product,liu2012movie,abulaish2009feature}.  While platforms can manipulate the display of user data, most have an incentive to be seen as truthful.}
The principal can incentivize agents to pull particular arms by
offering rewards for the specific arm.
The principal's goal is to keep the cumulative regret across all
agents small, while incurring only small total payments.
Our main theorem can be stated informally as follows:

@@

\begin{theorem} \label{thm:main-intro}
Assume that for each arm, at least a constant fraction \MinProb
of the population likes this arm best,
and let \TieDensity be a measure of the ``density of near-ties''
between agents' arm preferences
(in a sense made precise in Section~\ref{subsec:discrete}).
There is a policy that achieves expected 
cumulative regret $O (\ARMNUM \e^{2/\MinProb} + \dkedit{\TieDensity} \ARMNUM \log^3(T))$,
using expected cumulative payments of $O(\ARMNUM^2 \e^{2/\MinProb})$.
In particular, when agents who are close to tied between two arms have measure $0$,
both the expected regret and expected payment are bounded by constants
(with respect to $T$). 
\end{theorem}

\dkcomment{Added the \TieDensity term in the statement.}

### 2018-06-05 lower_bound.tex
significant edits throughout

We saw in Theorem~\ref{rst:discrete} that 
when agent preferences for their best arm are sufficiently clear,
in the sense that $\TieDensity = 0$, the regret of
Algorithm~\ref{alg:basic-incentivizing} is bounded by a constant.
One may conjecture that this should hold more generally,
in that the regret of the (fewer and fewer) agents on the boundary
between close arms should go to 0, while their fraction also goes to 0.
In this section, we establish a lower bound,
showing that even in very simple settings, a (logarithmic) dependence
on $T$ is typicaly unavoidable for \emph{any} incentivization strategy.

We consider an instance with two arms,
whose attribute vectors are $(0,0)$ and $(0,1)$, respectively.
Agents types are distributed uniformly on the (edge of the)
two-dimensional unit square%
\footnote{While our model technically assumed that all agent coordinates are
non-negative, we could simply shift the unit square.
The present choice is solely for ease of notation.}  
$\Set{\AgV}{\max(\Ag{1}, \Ag{2}) = 1}$,
with density $\frac{1}{8}$.
Because $\AgV \cdot (0,0) = 0$ and $\AgV \cdot (0,1) = \Ag{2}$,
the best choice for agent \AgV is arm $(0,1)$ iff $\Ag{2} > 0$,
i.e., the top half of the unit circle square the arm $(0,1)$,
and the bottom half prefers the arm $(0,0)$.

>>>>

We saw in Theorem~\ref{rst:discrete} that 
when agent preferences for their best arm are sufficiently clear,
in the sense that $\TieDensity = 0$, the regret of
Algorithm~\ref{alg:basic-incentivizing} is bounded by a constant.
One may conjecture that this should hold more generally,
in that the regret of the (fewer and fewer) agents on the boundary
between close arms should go to 0, while their fraction also goes to 0.
In this section, we establish a lower bound,
showing that even in very simple settings, a (logarithmic) dependence
on $T$ is typically unavoidable for \emph{any} incentivization strategy.

We consider an instance with two arms,
whose attribute vectors are $(0,0)$ and $(0,1)$, respectively.
Agent types are distributed uniformly on the (edge of the)
two-dimensional unit square%
\footnote{While our model technically assumed that all agent coordinates are
non-negative, we could simply shift the unit square.
The present choice is solely for ease of notation.}  
$\Set{\AgV}{\max(|\Ag{1}|, |\Ag{2}|) = 1}$,
with density $\frac{1}{8}$.
\dkcomment{I added the absolute values. Seemed wrong without, but
  please check!}
Because $\AgV \cdot (0,0) = 0$ and $\AgV \cdot (0,1) = \Ag{2}$,
the best choice for agent \AgV is arm $(0,1)$ iff $\Ag{2} > 0$,
i.e., the top half of the unit square prefers the arm $(0,1)$,
and the bottom half prefers the arm $(0,0)$.

### 2018-06-05 payment-proof.tex
significant edits throughout

To reason about such estimation errors formally, we define
\emph{envelopes} of sample paths.
A sample path \SP captures all the random events affecting the
algorithm, i.e., the random draws \AgentV{t} of agents and the
noise \NoiseV[t] in the draws the pulled arms,
with an infinite time horizon.

@@

>>>>

To reason about such estimation errors formally, we define
\emph{envelopes} of sample paths.
A sample path \SP captures all the random events affecting the
algorithm, i.e., the random draws \AgentV{t} of agents and the
noise \NoiseV[t] in the draws of the pulled arms,
with an infinite time horizon.

@@

\dkcomment{Did we forget a sum over \SP here?}

### 2018-06-05 problem_setting.tex
significant edits throughout

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
\pfedit{(We are also free to create new attributes that are nonlinear transformations of existing attributes.)}

For each time $t$ and arm $i$, let \NumPull{t}{i} be the number of
times that arm $i$ has been pulled (strictly) before time $t$.
An agent at time $t$ estimates arm $i$'s attribute vector as the
average of vectors observed during the arm's past pulls:
$\ArmEV{t}{i} = \frac{1}{\NumPull{t}{i}+1} \cdot
(\ArmEV{0}{i} + \sum_{t'<t: \Pull{t'} = i} \ObsV{t'})$;
here, \ArmEV{i}{0} is a single draw $\ArmV{i} + \NoiseV$ for arm $i$.
(In other words, we assume each arm is pulled once for free at time 0.)

@@

\pfedit{
We assume above that the distribution $\AgentDist$ over preference vectors is known.  In practice, one would estimate this from agents' selections and the attribute estimates and offered payments on which they were based.  While we do not show it, we hypothesize that our regret guarantees only change by constants as long as our estimate of the conditional probability of pulling an arm is correct within a constant factor (our algorithm depends on this quantity).
}

@@

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV,
breaking ties arbitrarily (but consistently).

@@

Let $\MinProb = \min_{i} \Prob[\AgentDist]{\Best{\AgV} = i}$
denote the minimum (over all arms) fraction of users that prefer any
particular arm.

@@

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exists a constant \TieDensity such that for all $z \in \R^+$,
%for all pairs $i,i'$ of arms and 
$\AlmostTied{z} \leq \TieDensity \cdot z$.
\end{assumption}

We emphasize that this assumption implies ties happen with probability $0$.

\paragraph{Role of parameters:}
Our problem setting is characterized by many 
parameters.
Of these, we consider \ARMNUM, $\MinProb \leq 1/\ARMNUM$ and $T$ to be
the key parameters, while we consider $\sigma, \Diam, d, \TieDensity$ to be
constant.
We track all parameters through most of our proofs,
but report final results in terms of only the three key parameters,
except where we illustrate a specific point.

>>>>

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
\pfedit{(We are also free to create new attributes that are nonlinear transformations of existing attributes.)}
\dkcomment{I have no idea what this sentence refers to.}

For each time $t$ and arm $i$, let \NumPull{t}{i} be the number of
times that arm $i$ has been pulled (strictly) before time $t$.
An agent at time $t$ estimates arm $i$'s attribute vector as the
average of vectors observed during the arm's past pulls:
$\ArmEV{t}{i} = \frac{1}{\NumPull{t}{i}+1} \cdot
(\ArmEV{0}{i} + \sum_{t'<t: \Pull{t'} = i} \ObsV{t'})$;
here, \ArmEV{0}{i} is a single draw $\ArmV{i} + \NoiseV$ for arm $i$.
(In other words, we assume each arm is pulled once for free at time 0.)
\dkedit{For a justification of the assumption that agents can observe
the actual noisy vectors \ObsV{t'} of past pulls, see
Section~\ref{sec:introduction}.}

@@

We assume above that the distribution $\AgentDist$ over preference
vectors is known to the principal.
In practice, a principal would estimate this distribution from agents'
selections and the attribute estimates and offered payments on which
they were based.
While we do not show it, we hypothesize that our regret guarantees
only change by constants as long as the principal's estimate of the
conditional probability of pulling an arm is correct to within a
constant factor.
\dkdelete{(our algorithm depends on this quantity).}

@@

To support the formulation of our results and the analysis,
we define the following additional notation.
We let
$\Best{\AgV} \in \argmax_i \AgV \cdot \ArmV{i}$
and
$ \Second{\AgV} \in \argmax_{i \neq \Best{\AgV}} \AgV \cdot \ArmV{i}$
denote the (indices of) the best and second-best arms for an agent
with attribute vector \AgV,
breaking ties arbitrarily (but consistently).
\dkedit{Notice that based on Assumption~\ref{A1} below,
  \Best{\AgV} and \Second{\AgV} are unique with probability 1.}

@@

Let $\MinProb = \min_{i} \Prob[\AgV \sim \AgentDist]{\Best{\AgV} = i}$
denote the minimum (over all arms) fraction of users that prefer any
particular arm.

@@

\begin{assumption}[Not too many near-ties] \label{A1}
Near-ties have vanishing probability; 
that is, there exist constants
\dkedit{$\TieDensity, \TieCutoff \geq \frac{\MinProb}{2 \TieDensity}$
such that 
$\AlmostTied{z} \leq \TieDensity \cdot z$ for all $z \leq \TieCutoff$.}
\end{assumption}

\dkedit{
The requirement that $\TieCutoff \geq \frac{\MinProb}{2 \TieDensity}$
is not restrictive.
Because \TieDensity and \TieCutoff are both constants,
we could simply increase \TieDensity.
Perhaps even more simply, $\MinProb \leq 1/\ARMNUM$,
while \TieCutoff is a constant,
so an alternative is to decrease \MinProb by a suitable constant factor.

Assumption~\ref{A1} implies that ties happen with probability $0$.
One special case of interest we discuss below is
$\TieDensity = 0$. This case arises when there is a cutoff
\TieCutoff such that only a measure-zero set of agents has two best arms
within utility \TieCutoff of each other.
In particular, $\TieDensity = 0$ happens whenever \AgentDist is discrete,
i.e., supported on a finite set of types.}

\paragraph{Roles of parameters:}
Our problem setting is characterized by many parameters.
Of these, we consider \ARMNUM, $\MinProb \leq 1/\ARMNUM$ and $T$ to be
the key parameters, while we consider
$\sigma, \Diam, d, \TieDensity, \TieCutoff$ to be constant.
We track all parameters through most of our proofs,
but report final results in terms of only the three key parameters,
except where we illustrate a specific point.
\dkedit{In particular, we consider $d$ a constant, because users
typically evaluate products (or restaurants) by a small number of
relevant features.}

### 2018-06-05 result_discussion.tex
significant edits throughout

When $\TieDensity = 0$, the bound of Theorem~\ref{rst:regret} is
constant in $T$;
thus, the algorithm achieves constant regret for constant payment. 
The case $\TieDensity = 0$ arises, for instance, in the case of a
discrete distribution for agents.
In fact, in the case of a discrete distribution of agents,
it is possible to modify the algorithm to reduce the dependence on
\MinProb from exponential to polynomial.
Theorem~\ref{rst:discrete} (given later) states that the modified
algorithm achieves expected regret
$O \left(\frac{\ARMNUM}{\MinProb^4} \right)$
with expected payments of
$O \left(\frac{\ARMNUM^2}{\MinProb^{5/2}} \right)$.

>>>>

When $\TieDensity = 0$, the bound of Theorem~\ref{rst:regret} is
constant in $T$;
thus, the algorithm achieves constant regret using constant payment. 
\dkedit{As discussed in Section~\ref{sec:prob},
the case $\TieDensity = 0$ arises, for instance, for
discrete agent distributions.}
In fact, in the case of a discrete agent distribution,
it is possible to modify the algorithm to \dkedit{also}
reduce the dependence on \MinProb from exponential to polynomial.
Theorem~\ref{rst:discrete} (given later) states that the modified
algorithm achieves expected regret
$O \left(\frac{\ARMNUM}{\MinProb^4} \right)$
with expected payments of
$O \left(\frac{\ARMNUM^2}{\MinProb^{5/2}} \right)$.
